---
title: "UFO_analysis"
author: "Surekha Peethambaram Muralidhar, Mrudula Nimmala, Sakshi Gundawar"
date: "2023-11-16"
output:
  html_document:
    keep_md: yes
    self_contained: yes
    pandoc_args: "--webtex"
  word_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
if (!knitr::is_html_output()) {
  knitr::opts_chunk$set(webshot = TRUE)
}
```

```{r}
library(dplyr)
library(ggplot2)
library(openintro)
library(knitr)
library(gtsummary)
library(tidyr)
library(readr)
library(tidyverse)
library(sf)
library(mapview)


ufo_sightings_transformed <- read_csv("ufo-sightings-transformed.csv")
```
## Corece Into Factors
```{r}
 ufo_sightings_transformed <- ufo_sightings_transformed %>% 
      mutate(Season = factor(Season),
             Country_Code = factor(Country_Code),
             Country = factor(Country),
             UFO_shape = factor(UFO_shape),
             Country = factor(Country),
             Year = factor(Year),
             Hour = factor(Hour),
             Region = factor(Region)
      )
```

## Summary Stats
```{r}

library(dplyr)
library(gtsummary)


# Create the summary table
UFO_summary_table <- ufo_sightings_transformed %>%
  select(length_of_encounter_seconds, Season,Month) %>%
  gtsummary::tbl_summary(
    statistic = list(
      all_continuous() ~ "{mean} ({sd})",
      all_categorical() ~ "{n} ({p}%)"),
    )
# Print the summary table
UFO_summary_table

```

## Including Plots

You can also embed plots, for example:

```{r}

# Install and load necessary packages
if (!requireNamespace("ggplot2", quietly = TRUE)) {
  install.packages("ggplot2")
}
library(ggplot2)

# Create a bar plot with reduced width
ggplot(ufo_sightings_transformed, aes(x = Season)) +
  geom_bar(fill = "skyblue", color = "darkblue", width = 0.5) +
  labs(title = "Bar Plot For UFO Sightings appeared in Different Seasons", x = "Season", y = "Count") +
  theme_minimal()

```

```{r}

ufo_sightings_transformed <- na.omit(ufo_sightings_transformed)

```

## 1.Write the hypotheses

$H_0:$ The average UFO sighting in California Region is equal to that of the average UFO sightings in England Region.


$\mu_{cal}=\mu_{eng}$


$H_A:$ The average UFO sighting in California Region is greater than that of the average UFO sightings in England Region.
            
$\mu_{cal} > \mu_{eng}$

## 2.Check Conditions

```{r}
california_england_data <- ufo_sightings_transformed %>%
  filter(Region %in% c("California", "England")) %>%
  filter(!is.na(Region)) %>%
  filter(length_of_encounter_seconds >= 0.1, length_of_encounter_seconds <= 15000, is.finite(length_of_encounter_seconds))

# Sample 1000 rows including both regions
sampled_data <- california_england_data %>%
  group_by(Region) %>%
  sample_n(500) %>%
  ungroup()

sampled_data %>% 
  ggplot(aes(length_of_encounter_seconds, fill = Region)) +
  geom_histogram(binwidth = 200, col = "white", show.legend = FALSE) +
  facet_wrap(~ Region) +
  labs(title = "Length of Encounter Seconds vs. Region") +
  scale_x_continuous(breaks = seq(0, 5000, by = 500), limits = c(0, 5000))

```

## 3. Test Statistic

```{r}

California_ufo_sightings <- ufo_sightings_transformed %>%
  filter(Region == "California", !is.na(length_of_encounter_seconds)) %>%
  select(length_of_encounter_seconds) %>%
  pull()

england_ufo_sightings <- ufo_sightings_transformed %>%
  filter(Region == "England", !is.na(length_of_encounter_seconds)) %>%
  select(length_of_encounter_seconds) %>%
  pull()



n1 <- length(California_ufo_sightings)
n2 <- length(england_ufo_sightings)

```

$n_1$ = `r n1`

$n_2$ = `r n2`

```{r}
xbar1 <- mean(California_ufo_sightings)
xbar2 <- mean(england_ufo_sightings)

s1 <- sd(California_ufo_sightings)
s2 <- sd(england_ufo_sightings)

stat <- xbar1 - xbar2
null_value <- 0
se <- sqrt(s1^2/n1 + s2^2/n2)

df <- min(c(n1, n2)) - 1

t_stat <- (stat - null_value) / se 

```

$\mu_{cal}$=`r round(xbar1,4)`

$\mu_{eng}$=`r round(xbar2,4)`

$t_{stat}$ = `r round(t_stat, 4)`

## 4.P-Value

```{r}
p_val <- pt(t_stat, df = df, lower.tail = FALSE)

```

$p-value$ = `r round(p_val, 4)`

Decision:Fail to reject $H_0$

Conclusion: We do not enough evidence that the difference in average of UFO sightings in California is greater than that average of UFO sightings in England.

```{r}
t.test(California_ufo_sightings, england_ufo_sightings, alternative = "greater",
       conf.level = 0.95)
```

# MapView For Califorina

```{r, results='asis', echo=FALSE}
if (knitr::is_html_output() || knitr::is_latex_output()) {
  ufo_cali <- ufo_sightings_transformed %>% 
    filter(Region == "California")
  mapview(ufo_cali, xcol = "longitude", ycol = "latitude", crs = 4269, grid = FALSE)
}
```



# MapView Englnd Region

```{r, results='asis', echo=FALSE}
if (knitr::is_html_output() || knitr::is_latex_output()) {
  ufo_england <- ufo_sightings_transformed %>% 
    filter(Region == "England")
  mapview(ufo_england, xcol = "longitude", ycol = "latitude", crs = 4269, grid = FALSE)
}
```


# WORLD-MAP

```{r, results='asis', echo=FALSE}

# Basic map
world <- map_data("world")

if (knitr::is_html_output() || knitr::is_latex_output()) {
  ggplot() +
    geom_polygon(data = world, aes(x = long, y = lat, group = group), fill = "grey1", color = "black") +
    geom_point(data = ufo_sightings_transformed, aes(x = longitude, y = latitude), color = "red3", size = 1) +
    labs(title = "UFO Sightings Map", subtitle = "Cities with UFO sightings") +
    theme_minimal()
}
```


